Survival Probability Difference M-F

### 2025-01-31 code/R/Survival_Probability_Difference_MF.Rmd
update figure

# Determine number of posterior draws (M) and time periods (T)
M <- nrow(phi_draws)
T <- ncol(eps_phi_matrix)

@@

# Define a helper function to compute summary statistics
summarize_draws <- function(x) {
  data.frame(
    mean          = mean(x),
    median        = median(x),
    sd            = sd(x),
    lower95       = quantile(x, 0.025),
    upper95       = quantile(x, 0.975),
    prob_positive = mean(x > 0)   # Proportion of draws where male > female
  )
}

# Apply the summary function to each time period's difference
time_summaries_list <- lapply(1:T, function(t) {
  summarize_draws(diff_survival[, t])
})

# Combine the summaries into a single data frame
time_summaries_df <- bind_rows(time_summaries_list, .id = "time_index") %>%
  mutate(time_period = paste0("time_", time_index)) %>%
  select(time_period, everything())

# View the summaries
print(time_summaries_df)

@@

# plot

# Ensure 'time_period' is an ordered factor (if not already done)
time_summaries_df <- time_summaries_df %>%
  mutate(time_period = factor(time_period, 
                              levels = paste0("time_", 1:T),  # Define the correct chronological order
                              ordered = TRUE))

# Verify the ordering
print(levels(time_summaries_df$time_period))

# Create the plot
difference_plot <- ggplot(time_summaries_df, aes(x = time_period, y = mean)) +
  geom_point(color = "blue", size = 2) +  # Increased point size for better visibility
  geom_errorbar(aes(ymin = lower95, ymax = upper95), width = 0.2, color = "blue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Difference in Survival Probabilities (Male - Female) Over Time",
    x = "Time Period",
    y = "Difference in Survival Probability"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Tilt x-axis labels for readability
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")  # Center and style the title
  ) +
  scale_x_discrete(limits = paste0("time_", 1:T))  # Ensure correct order on x-axis

@@

# Save the plot to a file
ggsave("/Users/johnimperato/Desktop/Mossy-Pond_companion/difference_in_survival_probabilities.png", plot = difference_plot, width = 10, height = 6, dpi = 300)

>>>>

# Determine number of posterior draws (M) and time periods (T)
M <- nrow(phi_draws)
T <- (ncol(eps_phi_matrix))

@@

# function to compute summary statistics
summarize_draws <- function(x) {
  data.frame(
    mean          = mean(x),
    median        = median(x),
    sd            = sd(x),
    lower95       = quantile(x, 0.025),
    upper95       = quantile(x, 0.975),
    prob_positive = mean(x > 0)   # Proportion of draws where male > female
  )
}

# Apply the summary function to each time period's difference
time_summaries_list <- lapply(2:(T-1), function(t) {
  summarize_draws(diff_survival[, t])
})

# Convert from factor to numeric via character
time_summaries_df <- time_summaries_df %>%
  mutate(time_index = as.numeric(as.character(time_index)))

# Combine the summaries into a single data frame
time_summaries_df <- bind_rows(time_summaries_list, .id = "time_index") %>%
  mutate(time_index = as.numeric(time_index)) %>%
  mutate(primary_period = paste("PPs_", time_index, "-", time_index + 1)) %>%
  select(primary_period, everything())

# View the summaries# Viewpaste0() the summaries
print(time_summaries_df)

@@

# Ensure 'primary_period' is an ordered factor to maintain chronological order in the plot
time_summaries_df <- time_summaries_df %>%
  mutate(primary_period = factor(primary_period, 
                                 levels = unique(primary_period), 
                                 ordered = TRUE))

# Verify the ordering
print(levels(time_summaries_df$primary_period))

# Create the plot
difference_plot <- ggplot(time_summaries_df, aes(x = primary_period, y = mean)) +
  geom_point(color = "blue", size = 2) +  # Increased point size for better visibility
  geom_errorbar(aes(ymin = lower95, ymax = upper95), width = 0.2, color = "blue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Difference in Survival Probabilities (M - F) Between Primary Periods",
    x = "Primary Period",
    y = "Difference in Survival Probability"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Tilt x-axis labels for readability
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")  # Center and style the title
  ) +
  # Correct the x-axis limits to match 'primary_period' labels
  scale_x_discrete(limits = levels(time_summaries_df$primary_period))

@@

# Save the plot to a file
ggsave("/Users/johnimperato/Desktop/Mossy-Pond_companion/difference_in_survival_probabilities.png", 
       plot = difference_plot, width = 10, height = 6, dpi = 300)
```
